---
author: "YOUR NAME HERE"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE, error = TRUE, fig.height = 3)
library(tidyverse)
library(kableExtra)
source("../../scripts/viridis.R")
source("../../scripts/ggprob.R")
```

\newcommand{\E}{\mathsf{E}}
\newcommand{\Var}{\mathsf{Var}}
\renewcommand{\prob}{\mathsf{P}}


## Assignment 7

#### Due Sunday, April 2, 11:59 PM CT

### Preliminaries

- Directories
    - COURSE/homework/
    - COURSE/homework/hw07/
    - COURSE/data/
    - COURSE/scripts/
- Files
  - COURSE/homework/hw07/hw07.Rmd
  - COURSE/scripts/viridis.R
  - COURSE/scripts/ggprob.R

### Aims

- Practice probability and the binomial distribution

### Problems

  1. A discrete random variable $X$ has possible values 0, 1, 2, 3, 4, 5, 6, 7, 8 with the following partial probability distribution.
The missing probabilities $\prob(X=7)$ and $\prob(X=8)$ are equal to each other.

```{r, echo = FALSE}
set.seed(20221021)
x = 0:8
p = rnorm(7,5,2.1)
p = sort(round(p / sum(p) * 0.84, 2))
p1 = tibble(x = x, p = c(p, rep(NA,2)))

p1_tab = p1 %>% 
  mutate(x = str_c(x)) %>% 
  pivot_wider(names_from = x, values_from = p) %>% 
  mutate(x = "P(X=x)", .before = `0`) 

p1_tab %>% 
  kable() %>% 
  kable_styling(position = "center", full_width = FALSE)
```

- Put the values `x = 0:8` and the tabled probabilities `p` into a data frame (use `tibble()`).

```{r}
q1_df = tibble(x = 0:8,
               p = c(0.04, 0.06, 0.11,	0.12,	0.14,	0.17,	0.2,	NA,	NA))
```

- Write code to calculate the missing values for $\prob(X=7) = \prob(X = 8)$ and replace the missing values with these calculated values into your data frame.

```{r}
missing_prob = (1 - sum(q1_df$p, na.rm = TRUE))/2
q1_df$p[is.na(q1_df$p)] = missing_prob
missing_prob
```

- Verify that the sum of all probabilities is 1.

```{r}
print(sum(q1_df$p))
```

- Calculate the mean, variance, and standard deviation of this distribution.

```{r}
#Calculate the mean
mean = sum(q1_df$x * q1_df$p, na.rm = TRUE)
print(mean)

#Calculate the variance
var = sum(((q1_df$x - mean)^2) * q1_df$p, na.rm = TRUE)
print(var)

#Calculate the standard deviation
sigma = sqrt(var)
print(sigma)

```






  2. Suppose you have a random variable $X \sim \text{Binomial}(120, 0.2)$.

- 2a. Calculate and report the mean and standard deviation of the distribution.

```{r}
n = 120
p = 0.2

mean = n*p
print(mean)

sd = sqrt(n*p*(1-p))
print(sd)
```

- 2b. Calculate and report the probability that the random variable is exactly equal to 20, $\prob(X = 20)$.

```{r}
px_20 = dbinom(20, n, p)
print(px_20)
```

- 2c. Calculate and report the probability that $X$ equals 20 or more, $\prob(X \ge 20)$.

```{r}
px_20_or_more = pbinom(19, n, p, lower.tail = FALSE)
print(px_20_or_more)

```

- 2d. Draw a graph of the binomial distribution with vertical lines at the mean (red and dashed), and 1, 2, and 3 standard deviations above and below the mean (red and dotted).


```{r}
x = 0:n
probs = dbinom(x, n, p)
data = tibble(x, probs)

ggplot(data,aes(x=x, y = probs))+
  geom_col(fill = "lightblue")+
  geom_vline(xintercept = mean, linetype = "dashed", color = "red") +
  geom_vline(xintercept = mean - sd, linetype = "dashed", color = "red") +
  geom_vline(xintercept = mean - 2*sd, linetype = "dashed", color = "red") +
  geom_vline(xintercept = mean - 3*sd, linetype = "dashed", color = "red") +
  geom_vline(xintercept = mean + sd, linetype = "dashed", color = "red") +
  geom_vline(xintercept = mean + 2*sd, linetype = "dashed", color = "red") +
  geom_vline(xintercept = mean + 3*sd, linetype = "dashed", color = "red") +
  ggtitle("Binomal Distribution")+
  xlab("X")+
  ylab("Probability")+
  theme_minimal()
```





  3. What is the probability that $X$ from Problem 2 is within one, two, and three standard deviations of the mean? Round each probability to four decimal places.

```{r}
#Probability within one standard deviation of the mean
p1 = round(pnorm(mean + sd, mean, sd) - pnorm(mean - sd, mean, sd), 4)
p1

#Probability within two standard deviations of the mean
p2 = round(pnorm(mean + 2*sd, mean, sd) - pnorm(mean - 2*sd, mean, sd), 4)
p2
#Probability within three standard deviations of the mean
p3 = round(pnorm(mean + 3*sd, mean, sd) - pnorm(mean - 3*sd, mean, sd), 4)
p3

print(paste0("The probability that $X$ from Problem 2 is within one standard deviation is ", p1))
print(paste0("The probability that $X$ from Problem 2 is within two standard deviations is ", p2))
print(paste0("The probability that $X$ from Problem 2 is within three standard deviations is ", p3))
```






  4. The following code makes a graph of the cumulative distribution function for the $\text{Binomial}(7, 0.3)$ distribution, that is $F(x) = \prob(X \le x)$. This is the function which is calculated by the base R function `pbinom()`.

- Modify the code to add a horizontal red dotted line at 0.4.


```{r}
prob4 = tibble(
  x = seq(-1,8,1),
  p = dbinom(x, 7, 0.3),
  cdf = pbinom(x, 7, 0.3))

ggplot(prob4, aes(x = x, y = cdf )) +
  geom_step(color = "blue") +
  geom_hline(yintercept = 0.4, linetype = "dotted", color = "red") +
  scale_x_continuous(breaks = 0:7) +
  xlab("x") +
  ylab("Probability") +
  ggtitle("Binomial(7, 0.3) distribution CDF") +
  theme_minimal()
```

- Calculate the 0.4 quantile of the $\text{Binomial}(7, 0.3)$

```{r}
quantile_40 = qbinom(0.4, 7, 0.3)
quantile_40
```

- Explain how the value of the quantile is related to the red dotted line which you added to the plot.

> [The 0.4 quantile of the Binomial(7, 0.3) distribution is the smallest integer value of x such that the cumulative probability P(X <= x) is greater than or equal to 0.4. We have calculated this to be 2.
  The red dotted line we added to the plot represents the cumulative probability. In other words, the calculated value has the value of x at which the blue step function in the plot crosses the red dotted line. From the graph, the step funtrion crosses the dotted line at x = 2. Therefore, the 0.4 quantile of the Binomial(7, 0.3) distribution is where the red dotted line in the plot crosses the blue step function in the plot.]

- Use the graph to determine the 0.75 quantile of this distribution.
Then verify your observation with a calculation.

> [From the graph, the value x at 0.75 quantile is 3.]

```{r}
#Check
quantile_75 = qbinom(0.75, 7, 0.3)
quantile_75
```





  5. In a large population of likely voters, the proportion of voters who plan to vote for candidate A is $p = 0.52$. In a random sample of 600 likely voters, what is the probability that the proportion of voters who plan to vote for candidate A is less than 50% (the number of such voters is less than 300)?
  
```{r}
n = 600
p = 0.52
prob = pbinom(299, n, p)


prob

```

```{r}
# Set the parameters of the binomial distribution
p <- 0.52
n <- 600

# Calculate the mean and standard deviation of the normal distribution
mu <- n * p
sigma <- sqrt(n * p * (1 - p))

# Calculate the z-score for X = 0.5
z <- (0.5 - mu) / sigma

# Calculate the probability P(X < 0.5)
p <- pnorm(z)
p

```




  6. Basketball player 1 makes 80% of their free throws and basketball players 2 makes 70% of their free throws. Assume all of their free throws attempts are independent. If each player takes ten free throws and $X$ is the total number of free throws they make, does $X$ have a binomial distribution? If yes, specify the parameters $n$ and $p$. If not, explain why not.
  
> [X does not have a binomal distribution., but rather the sum of two different binomal distributions. 
An individual player's success rate of free throws can be modeled as a binomial distribution. For player 1, n1 = 10 (number of free throws), and p1 = 0.8 (probability of success)]




  7. What is the smallest number of coin tosses $n$ of a fair coin so that the chance of getting 50 or more heads exceeds 90%?  (You may need to do some trial and error. Alternatively, think of creating a data frame with a column `n` and a column for the probability and using this to find the answer.)

```{r}

# create a function to calculate the probability of getting 50 or more heads
prob_50_or_more_heads <- function(n) {
  p <- sum(dbinom(50:n, n, 0.5)) # probability of getting 50 or more heads
  return(p)
}

# create a data frame with number of tosses and the corresponding probability
df <- tibble(n = 1:1000,
                 p = sapply(1:1000, prob_50_or_more_heads))

# find the smallest number of tosses for which the probability exceeds 90%
smallest_n <- min(df$n[df$p > 0.9])
smallest_n

```

```{r}
# Initialize variables
p <- 0.5


# Create a vector of values of n from 1 to 1000
n_vec <- 1:1000

# Calculate the probability of getting 50 or more heads for each value of n
prob_vec <- 1 - pbinom(49, n_vec, p)

# Find the smallest value of n for which the probability exceeds 0.9
n_smallest <- min(n_vec[prob_vec > 0.9])

# Print the result
print(paste0("The smallest number of coin tosses is: ", n_smallest))

```



  8. The random variable $X$ has the $\text{Binomial}(100, 0.2)$ distribution.
  
- Find an integer $a$ so that $P(X \le a) \ge 0.5$ and $P(X \ge a) \ge 0.5$.
- Show the values of $a$, $P(X \le a)$, and $P(X \ge a)$.

```{r}
# define the parameters of the Binomial distribution


a = qbinom(0.5, 100, 0.2, lower.tail = FALSE)
p = pbinom(a, 100, 0.2)
#p_complement = 1 - p
#p_complement

p_complement = 1 - pbinom(a - 1, 100, 0.2)
p_complement

print(paste0("a = ", a))
print(paste0("P(X<=a) = ", p))
print(paste0("P(X>=a) = ", p_complement))
 
```


  


